---
title: "AhmedElsaeyed-TidyVerseCreate"
author: "Ahmed Elsaeyed"
date: "4/9/2022"
output:
  html_document:
    df_print: paged
    fig_caption: true
    theme: cosmo

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message= FALSE)
library(dplyr) 
library(stringr)
library(lubridate)
library(ggplot2)
```


## Data Import

To begin I will import my data, which I downloaded from: https://www.kaggle.com/datasets/datota/fruit-and-vegatable-prices-in-uk-2017-2022?resource=download. This is a dataset that tracks the average prices of fruits on a weekly basis in the UK. My goal is to plot the trend in prices for apples over the years to see when they tend to be cheapest and presumably in-season and tasty.
```{r}
groc_prices = read.csv("https://raw.githubusercontent.com/aelsaeyed/Data607/main/fruitvegprices-2017_2022.csv")

head (groc_prices, 10)
```
## Filter
To achieve this I will use an array of tools from Tidyverse, mainly dplyr and stringr. The first step is to filter for the data I want, mainly apples. 
```{r}
apples <- groc_prices %>%
  filter(item == "apples")

apples
```
## Side Quest
As a side-quest I decided to see which variety of apple is the most expensive on average. 

```{r side quest}
pricey_fruit <- apples %>% 
  group_by(variety) %>% 
  summarise(variety_price = max(mean(price))) %>% 
  arrange(desc(variety_price))

pricey_fruit
```

## Date Ordering

I decided to ignore the variety of apple as a relevant factor for now and chose to simply get the average price of all apples per day. I will then plot that data to see the trend. I had to do something about the date column first, which was in char format. Lubridate allows me to do this easily using the ymd function. I can now arrange by date descending.

```{r}
apples_clean_date <- apples %>% 
  mutate(`date` = ymd(date) ) %>% 
  arrange(desc(date))  # %>% 
  #select(date, price)

apples_clean_date
```

## Average Price Per Date
I now want just one price value per date to represent all the apples, and I will use the average price.

```{r average price per date}
weekly_apples <- apples_clean_date %>% 
  group_by(date) %>% 
  summarise( price_day = mean(price) ) %>% 
  arrange(desc(date))
weekly_apples
```
## Plotting
I can now create a graph of the prices as they change over the weeks:

```{r plot}
ggplot(data=weekly_apples, aes(x=date, y=price_day, group=1)) +
  geom_line()+
  geom_point() +
  scale_x_date(date_labels="%b %y",date_breaks  ="2 months") +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Price") + xlab("Date") +
  ggtitle("Price per Apple Weekly", "When is the cheapest time to keep the doctor away?") +
  geom_line(color="pink", size=1) + 
  geom_point(color="black", size=1)
```
The trend seems to be that the apples get more expensive in the summers, with a strange inversion in 2021.
