---
title: "TidyVerse Create"
author: "Al Haque"
editor: "Benson Toi"
date: "3/30/2022"
output: html_document
---
### TidyVerse Project
 Tidyverse is so commonly used in data science that it helps data scientists manipulate,tidy,transform and visualize data.The tidyverse contains a bunch of packages that helps beginners or expert manage data.In this project I will demonstrate the following package in the tidyverse this is the dyplr package.
 

### Dyplr 
  Dyplr is used in data science to help the user clean and manipulate data in a manner which they fill fitting. 
  
  
  ## Calling the Library and Reading the data 
 Here we imported the tidyverse packages which contains the package dplyr but you can also see other packages as well such as tibble,readr, and more.I also imported a data set from kaggle which contains information about the Disney movies. 

```{r}
library(tidyverse)
Disney_data <- read.csv("https://raw.githubusercontent.com/AldataSci/TidyVerse/main/disney_movies.csv",sep=",",header=TRUE)
```

### Looking at the data
  Let's take a look at the data and see what we can understand from it.We can see all the movie_title,the genre and the release date and its earnings as well. 
```{r}
head(Disney_data)
```

### Cleaning The Data
  We can see that there are a select few columns and we can remove from our data if we wish to
```{r}
## You can use the select function to select columns based on their names 

head(select(Disney_data,movie_title,release_date,genre,mpaa_rating,inflation_adjusted_gross))

## Alternatively you can also use the pipes dpylr to get the same results!! 

head(Disney_data %>%
  select(movie_title,release_date,genre,mpaa_rating,inflation_adjusted_gross))

## A much simpler way to get the columns you want without writing the whole columns name out is to write a minus sign next to the column you don't want.

head(Disney_data %>%
  select(-total_gross))

```

### Filtering the Data
  We can also filter out our data to make it more readable or simply more easy to understand. In our case we can see there is a lot of empty values under the ratings columns which simply fills the data with empty space and harder to understand.
  
```{r}
## The filter function helps us filter out the data I filtered out the observations in the rating columns that contains empty space or if the ratings says not rated

head(filter(Disney_data,mpaa_rating !="" & mpaa_rating !="Not Rated"))

## We can also use dplyr pipes to filter it out

head(Disney_data %>%
  filter(mpaa_rating !="" & mpaa_rating != "Not Rated"))

```
### Renaming Columns 
 We can also rename our columns in your dataset if the name doesn't make sense or if the name is ugly to you 
 
```{r}
## We can use base R rename function to simply rename our column names in our dataset.In this case we will rename our movie_title column to Movies 

head(names(Disney_data)[names(Disney_data) == "movie_title"] <- "Movies")

### Alternatively a simpler and faster way to do this is to use dyplr rename function where we will pipe the disney_data rename the columns we want and get a new dataset with the changed column names saved to it the new name you want to give goes to the left of =, in this case it is Movie_Titles I want to rename to while the right of = is the old name of the column Movies. 

head(Disney_data %>%
  rename(Movie_Titles=Movies))
```
### Combining Everything with the Pipe Operator
```{r}
## Here we can combine multiple pipes command as one we can put all the changes we did to the dataset into one R pipe command that is the awesome power of the pipe operator.. and we save our data in a dataframe called Clean Disney.. 

head(Clean_Disney <- Disney_data %>%
  rename(movies=Movies) %>%
  select(-total_gross) %>%
  filter(mpaa_rating !="" & mpaa_rating !="Not Rated"))

```
 
## Advanced Usages of the Pipes
 We can also use dyplr to re-organize our data and we can also perform analysis on our cleaned data now. 
 
```{r}
## The group-by command helps aggregate your data, aggregate data groups your data into a specific category, 
head(Clean_Disney %>%
  group_by(release_date))
  
```
```{r}
## The summarise function helps us perform numerical analysis on our data so using the max function we use the max value of the column. 
head(Clean_Disney %>%
  group_by(release_date) %>%
  summarise(max(inflation_adjusted_gross)))

```
```{r}
### We can also use the arrange function to rearrange our data from low-to high with the desc function arranging our data can help provide more additional insights to perform. 

head(Clean_Disney %>%
  arrange(desc(inflation_adjusted_gross)))
```
## Tibble

A tibble is a data frame that is designed for presentation.
```{r tibble}

Clean_Disney_new <- as_tibble(Clean_Disney)
print(Clean_Disney_new)

```
