---
title: "Tidyverse project"
author: "Jay Lee and Benson Toi"
date: "4/8/2022"
output: 
  html_document:
     theme: yeti
     toc: true
     toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project details

Our task here is to Create an Example.Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.


URL: https://raw.githubusercontent.com/fivethirtyeight/data/master/hate-crimes/hate_crimes.csv



# Tidyverse project - prepare the data

## Step 1 Load Library 

We need to load the Library first

Tidyverse has following packages

✓ ggplot2 3.3.5     ✓ purrr   0.3.4
✓ tibble  3.1.6     ✓ dplyr   1.0.7
✓ tidyr   1.1.4     ✓ stringr 1.4.0
✓ readr   2.1.2     ✓ forcats 0.5.1

```{r}
library(tidyverse)
```

## Step 2 Load data 

Then we load it into R

```{r data}
# load data

hate_url<- "https://raw.githubusercontent.com/fivethirtyeight/data/master/hate-crimes/hate_crimes.csv"
hate_url <-read.csv(hate_url)
head(hate_url)
```

## Step 3 add new column with calcalation

We replace the na to 0 and sum the hate crime cases together.

```{r}

haterate <-hate_url %>%
  replace(is.na(.), 0) %>%
  mutate(hate_rate_sum_per100k = rowSums(.[11:12]))

head(haterate)

```

# Tidyverse project - use funtion from Tidyverse

## glimpse data

We use glimpse to check the format of the data as well as how many columns and rows.

```{r}
glimpse(haterate)
```


## Select Data

We select the column that we need when some column is not needed to minimize the size of the data.

```{r}
haterate <- haterate %>%
select("state","median_household_income","share_unemployed_seasonal","hate_crimes_per_100k_splc","avg_hatecrimes_per_100k_fbi","hate_rate_sum_per100k")
head(haterate)


```

## create summary 

We create a summary to have a overview of the data, it is helpful to quick check if there is any outliner. 

```{r}
summary(haterate)
```

## ggplot

we try to visualize the count of unemployment and hate crimes cases. 

```{r}
ggplot(haterate, aes(x=hate_rate_sum_per100k)) + geom_histogram(bins = 30) 
ggplot(haterate, aes(x=share_unemployed_seasonal)) + geom_histogram(bins = 30)
```

## DD Extension

We'll start by adding some graphics-based libraries to produce additional visuals:

```{r}
library(ggthemes)
library(maps) 
library(ggpubr)
library(kable)
```

### Figures Adjustments

We'll start by visualizing unemployment rate by count:

```{r}

plt.hate.100k = ggplot(haterate, aes(x=hate_rate_sum_per100k)) + geom_histogram(bins = 30, color = "white", fill = "darkblue") + theme_minimal() + labs(x = "Hate Rate per 100K", y = "N")
plt.unemployment.100k = ggplot(haterate, aes(x=share_unemployed_seasonal)) + geom_histogram(bins = 30, color = "white", fill = "darkred")  + labs(x = "Unemployment Rate", y = "N") + theme_minimal()

ggarrange(plt.hate.100k, plt.unemployment.100k, ncol = 1)
```


### Correlation Plots

Ideally, we'd like to see if there's a correlation between these two values. Therefore, we'll start with a correlation plot. We'll also take it a step further by implementing a rough linear model:

```{r}
ggplot(data = haterate, aes(share_unemployed_seasonal, hate_crimes_per_100k_splc)) + geom_point() + geom_smooth() + theme_minimal() + labs(x = "Unemployment Rate", y = "Hate Crimes per 100k")

summary(glm(hate_crimes_per_100k_splc ~ share_unemployed_seasonal, data  = haterate))
```

### Choropleth

Lastly, we'll implement a visualization by region:

```{r}
MainStates <- map_data("state")

haterate = haterate %>% mutate(region = tolower(state))

MergedStates <- inner_join(MainStates, haterate, by = "region")


ggplot() + 
  geom_polygon( data=MergedStates, aes(x=long, y=lat, group=group, fill = median_household_income),
                color="white", size = 0.25) + theme_minimal() + scale_fill_viridis_b() + theme(legend.position = "bottom", axis.title.x = element_blank(), axis.title.y = element_blank(), panel.grid = element_blank(), axis.text = element_blank()) + labs(title = "Median Household Income by State")


ggplot() + 
  geom_polygon( data=MergedStates, aes(x=long, y=lat, group=group, fill = hate_rate_sum_per100k),
                color="white", size = 0.25) + theme_minimal() + scale_fill_viridis_b() + theme(legend.position = "bottom", axis.title.x = element_blank(), axis.title.y = element_blank(), panel.grid = element_blank(), axis.text = element_blank()) + labs(title = "Hate crimes per 100k by State")

```

