---
title: "Tidyverse project Extended"
author: "Originally: Jay Lee and Benson Toi  Extended By: Alex Moyse"
date: "4/8/2022"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<H1> Project details </H1>

Our task here is to Create an Example.Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset.


URL: https://raw.githubusercontent.com/fivethirtyeight/data/master/hate-crimes/hate_crimes.csv



<H1> Tidyverse project - prepare the data </H1>

<H2> Step 1 Load Library </H2> 


We need to load the Library first

Tidyverse has following packages

✓ ggplot2 3.3.5     ✓ purrr   0.3.4
✓ tibble  3.1.6     ✓ dplyr   1.0.7
✓ tidyr   1.1.4     ✓ stringr 1.4.0
✓ readr   2.1.2     ✓ forcats 0.5.1

```{r}
library(tidyverse)
```

<h2> Step 2 Load data </H2> 

Then we load it into R

```{r data}
# load data

hate_url<- "https://raw.githubusercontent.com/fivethirtyeight/data/master/hate-crimes/hate_crimes.csv"
hate_url <-read.csv(hate_url)
head(hate_url)
```

<H2>  Step 3 add new column with calcalation </H2> 

We replace the na to 0 and sum the hate crime cases together.

```{r}

haterate <-hate_url %>%
  replace(is.na(.), 0) %>%
  mutate(hate_rate_sum_per100k = rowSums(.[11:12]))

head(haterate)

```

# Tidyverse project - use funtion from Tidyverse

## glimpse data

We use glimpse to check the format of the data as well as how many columns and rows.

```{r}
glimpse(haterate)
```


## Select Data

<H2> We select the column that we need when some column is not needed to minimize the size of the data. </H2> 

```{r}
haterate <- haterate %>%
select("state","median_household_income","share_unemployed_seasonal","hate_crimes_per_100k_splc","avg_hatecrimes_per_100k_fbi","hate_rate_sum_per100k")
head(haterate)


```

<H2>  create summary </H2> 

We create a summary to have a overview of the data, it is helpful to quick check if there is any outliner. 

```{r}
summary(haterate)
```

## ggplot

we try to visualize the count of unemployment and hate crimes cases. 

```{r}
ggplot(haterate, aes(x=hate_rate_sum_per100k)) + geom_histogram(bins = 30) 
ggplot(haterate, aes(x=share_unemployed_seasonal)) + geom_histogram(bins = 30)
```


<H1> And now, the extension </H1> 

<H2> What is it? </H2>

Going to be brutally honest, my friend showed me this package for R and it piqued my interest, its named corrplot. It makes a wonderful chart to see relationships between things 

<H2> First things first, let's load it in </H2>

```{r}
library(corrr)
library(GGally)
```
<H2> Data Grab </H2>


This is pretty much the same as above!

```{r}
haterate <-hate_url %>%
  replace(is.na(.), 0) %>%
  mutate(hate_rate_sum_per100k = rowSums(.[11:12]))

```

<H2> And now the addition </H2>

Practically corr demontrates relationships between variables, as well as the degree of that relationship

```{r}
haterate_corr <- haterate %>%  select(-"state") %>% correlate()     %>%   rearrange() %>%  # rearrange by correlations
       shave() # Shave off the upper triangle for a clean result
fashion(haterate_corr)

ggcorr(haterate_corr, geom = "circle", hjust = 0.75, size = 5, layout.exp = 4)
```

<H1> Conclusion </H1>

I want to thank Jay Lee and Benson Toi for starting this out, it made a fantastic start to see the data as well as a chart, and I wanted to extend it to use more variables!

More info on the release can be found here:

https://www.tidyverse.org/blog/2020/12/corrr-0-4-3/

Alex 
