---
title: "DPLYR Vignette"
author: "Juan Falck"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# DPLYR

## Introduction
DPLYR is one of the Packages included in **TIDYVERSE**. The package simplifies the language used to select pieces of data set.

## Intialization
Let's load libraries we will need.

```{r}
#We will used readr to reac in a file to be used for the examples
library(readr)
library(dplyr)
```

If error, please ensure library is installed before by using the following command.  **install.packages("dplyr")** or **install.packages("readr")**

Let's load a data set we will use for testing DPLYR. We will load from **FiveThirtyEight** a dataset about **Drug Use by Age**

```{r}
drug_use <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/drug-use-by-age/drug-use-by-age.csv")
```


## Main functions

### Select
**SELECT** allows us to easily select **COLUMNS** from a dataset.

**USAGE**
select(data,column1,columns2..columnn)

Select a sing column

```{r}
drug_use %>%
  select(age) %>%
  head()
```

Select multiple columns.

```{r}
drug_use %>%
  select(age,`cocaine-use` ) %>%
  head()
```

You can select consecutive columns by using **:**
```{r}
drug_use %>%
  select(age:`marijuana-use` ) %>%
  head()
```


### Filter
**FILTER** allows of to select a subset of **ROWS** from the dataframe

**USAGE**
filter(data,expression)

Filtering by one criterion
```{r}
drug_use %>%
  filter(age>20)
```


Filtering by two criteria with an **AND**
```{r}
drug_use %>%
  filter(age>20 & `marijuana-use` <20)
```


Filtering by two criteria with an **OR**
```{r}
drug_use %>%
  filter(age>20 | `marijuana-use` <20)
```


Combining **SELECT** and **FILTER**
```{r}
drug_use %>%
  select(age,`alcohol-use`,`marijuana-use`) %>%
  filter(`alcohol-use`<20)
```


### Mutate
**MUTATE** allows us to add columns to our dataframes, which can be the result of operations between the existing ones.

**USAGE**
mutate(date,column_to_add)

Add a single columns based on another one (In this example some rows will have NA, but for purpose of example it would be ok)
```{r}
drug_use %>%
  mutate(age_months = as.numeric(age)*12) %>%
  select(age,age_months)
```



### Summarise

**SUMMARISE** creates a new data frame. It will have one (or more) rows for each combination of grouping variables; if there are no grouping variables, the output will have a single row summarising all observations in the input. It will contain one column for each grouping variable and one column for each of the summary statistics that you have specified.

**USAGE**
summarise(data, summary1, summar2, summaryn)

Output the number of records in the dataframe.

```{r}
drug_use %>%
  summarise(records = n())
```

Output the number of records by age-group
```{r}
drug_use %>%
  group_by(age) %>%
  summarise(records = n())
```

Sum a column under criteria.

```{r}
drug_use %>%
  filter(age>21) %>%
  summarise(mysum = sum(`cocaine-use`))
```
